---
title: "Pseudotime for clusters 16.0 and 16.1"
author: "Ramiro Lorenzo"
date: "April 8, 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
```


## Load library
```{r library,results='hide',message=FALSE}
library(monocle3,verbose = FALSE)
library(dplyr,verbose = FALSE)
```

## Read the data
```{r Data}
## Loads a sparse matrix RawCountsPseudotime
load("RawCountsPseudotime.rda")
dim(RawCountsPseudotime)
```

##Load cluster data
```{r}
## Loads a factor variabele ClusterPseudotime containing cluster identities
load("ClusterPseudotime.rda")
## Change the name of variable to remind
DataClusters <- ClusterPseudotime

table(DataClusters)
length(DataClusters)
```

##Convert DataClusters to a matrix format for input to Monocle
```{r}
DataCluster.ID <- matrix(as.numeric(levels(DataClusters))[DataClusters], ncol = 1)
rownames(DataCluster.ID) <- names(DataClusters)
colnames(DataCluster.ID) <- "Cluster.IDs"
DataCluster.ID[1:10,]
```

##Generate matrix of gene short names for Monocle
```{r}
geneNames <- matrix(rownames(RawCountsPseudotime), ncol = 1)
rownames(geneNames) <- rownames(RawCountsPseudotime)
colnames(geneNames) <- "gene_short_name"
head(geneNames)
```

##Initiate Monocle object
```{r}
cds <- new_cell_data_set(expression_data = RawCountsPseudotime,
                         cell_metadata = DataCluster.ID,
                         gene_metadata = geneNames)


cds <- cds[,names(ClusterPseudotime[ClusterPseudotime %in% c("16.0","16.1")])] 


colData(cds)

## Column Cluster.IDs contains the original DataCluster IDs
colData(cds)$Cluster.IDs <- factor(colData(cds)$Cluster.IDs)
colData(cds)
```

## Step 1: Normalize and pre-process the data
```{r}
cds <- preprocess_cds(cds, num_dim = 100)
plot_pc_variance_explained(cds)
cds <- preprocess_cds(cds, num_dim = 20)
```

##Step 2: Reduce the dimensions using UMAP
```{r}
cds <- reduce_dimension(cds,umap.min_dist = 0.1,cores = 8)
plot_cells(cds, color_cells_by = "Cluster.IDs", group_label_size = 4, cell_size = 0.5)
```

##Step 3: Cluster the cells
```{r}
# cds = cluster_cells(cds, resolution=0.02)
cds = cluster_cells(cds, resolution=0.02)
plot_cells(cds, cell_size = 0.5)

```

##Step 4: Learn a graph and order cells
```{r}
cds <- learn_graph(cds)
```
```{r}
## With Shiny
# cds <- order_cells(cds)
# #
# save(file = "cdsRoots.rda", cds) #Save object state after selection of roots

## Without Shiny
load("cdsRoots.rda") #Load object

print(cds@principal_graph_aux$UMAP$root_pr_nodes)
```
```{r}
plot_cells(cds)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "Cluster.IDs",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5)
```

##Plot the pseudotime graph without trajectory
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5,
           show_trajectory_graph = FALSE)
```

## Correlation Analysis
### Cluster 16
```{r,results='hide'}
cluster16.cellNames <- rownames(pData(cds))[pData(cds)$Cluster.IDs %in% c(16, 16.1)]

cds_16 <- cds[,cluster16.cellNames]

cds16_pg <- graph_test(cds_16, neighbor_graph="principal_graph", cores=4, verbose = F)

cds16_genes <- cds16_pg %>% 
  filter(q_value < 0.05) %>% 
  arrange(desc(morans_I)) %>% 
  select(gene_short_name)
```

```{r}
cds16_genes$gene_short_name
```


```{r}
# Plot a few genes
plot_cells(cds_16, genes=c("Y48G10A.6", "egl-44", "egl-46", "zag-1"),
           show_trajectory_graph=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           cell_size = 0.5)
```
```{r}
# Plot a few genes
plot_cells(cds_16, genes=c("alr-1", "sto-5", "dma-1", "flp-4"),
           show_trajectory_graph=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           cell_size = 0.5)
```
```{r}
# Plot a few genes
plot_cells(cds_16, genes=c("hbl-1", "lin-46","asic-1", "mec-10"),
           show_trajectory_graph=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           cell_size = 0.5)
```
```{r}
# Plot a few genes
plot_cells(cds_16, genes=c("mec-3", "egl-46", "dma-1", "lin-14"),
           show_trajectory_graph=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           cell_size = 0.5)
```

```{r}
# Plot a few genes
plot_cells(cds_16, genes=c("mec-3", "egl-46", "F49H12.4", "ser-2"),
           show_trajectory_graph=FALSE,
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           cell_size = 0.5)
```

#Clustering Genes by Pseudotemporal Expression Pattern
```{r}

cds_16_lineage_cds <- cds_16[rowData(cds_16)$gene_short_name %in% c("Y48G10A.6","dma-1", "lin-14", "egl-46"),]
plot_genes_in_pseudotime(cds_16_lineage_cds,
                         min_expr=0.05)
```
```{r}

cds_16_lineage_cds <- cds_16[rowData(cds_16)$gene_short_name %in% c("mec-3", "mec-10", "asic-1","F49H12.4", "ser-2"),]
plot_genes_in_pseudotime(cds_16_lineage_cds,
                         min_expr=0.05)
```