---
title: "Pseudotime for clusters 0 and 2"
author: "Ramiro Lorenzo"
date: "April 8, 2020"
output:
  pdf_document: default
  html_document: default
---

## Load library
```{r library,results='hide',message=FALSE}
library(monocle3,verbose = FALSE)
library(dplyr,verbose = FALSE)
```

## Read the data
```{r Data}
## Loads a sparse matrix RawCountsPseudotime
load("../RawCountsPseudotime.rda")
dim(RawCountsPseudotime)
```

##Load cluster data
```{r}
## Loads a factor variabele ClusterPseudotime containing cluster identities
load("../ClusterPseudotime.rda")
## Change the name of variable to remind
DataClusters <- ClusterPseudotime

table(DataClusters)
length(DataClusters)
```

##Convert DataClusters to a matrix format for input to Monocle
```{r}
DataCluster.ID <- matrix(as.character(levels(DataClusters))[DataClusters], ncol = 1)
rownames(DataCluster.ID) <- names(DataClusters)
colnames(DataCluster.ID) <- "Cluster.IDs"
DataCluster.ID[1:10,]
```

##Generate matrix of gene short names for Monocle
```{r}
geneNames <- matrix(rownames(RawCountsPseudotime), ncol = 1)
rownames(geneNames) <- rownames(RawCountsPseudotime)
colnames(geneNames) <- "gene_short_name"
head(geneNames)
```

##Initiate Monocle object
```{r}
cds <- new_cell_data_set(expression_data = RawCountsPseudotime,
                         cell_metadata = DataCluster.ID,
                         gene_metadata = geneNames)

cds <- cds[,names(ClusterPseudotime[ClusterPseudotime %in% c(
                                                                "0.0",
                                                                "0.1",
                                                                # "1.0",
                                                                # "1.1",
                                                                "2.0"
                                                                # "3.0",
                                                                # "3.1",
                                                                # "4.0",
                                                                # "4.1",
                                                                # "16.0",
                                                                # "16.1",
                                                                # "17.0",
                                                                # "17.1",
                                                                # "8.0.0",
                                                                # "8.0.1",
                                                                # "8.1",
                                                                # "8.2",
                                                                # "8.3",
                                                                # "10.0.0",
                                                                # "10.0.1_4",
                                                                # "10.0.2_3",
                                                                # "10.0.5",
                                                                # "10.1.0",
                                                                # "10.1.1",
                                                                # "10.2",
                                                                # "33.0",
                                                                # "42.0",
                                                                # "42.1"
                                                                # "56.0"
                                                                )])] #all data for trajectories



colData(cds)

## Column Cluster.IDs contains the original DataCluster IDs
colData(cds)$Cluster.IDs <- factor(colData(cds)$Cluster.IDs)
colData(cds)
```

## Step 1: Normalize and pre-process the data
```{r}
cds <- preprocess_cds(cds, num_dim = 100)
plot_pc_variance_explained(cds)
cds <- preprocess_cds(cds, num_dim = 10)

```

##Step 2: Reduce the dimensions using UMAP
```{r}
cds <- reduce_dimension(cds,cores = 8)
plot_cells(cds, color_cells_by = "Cluster.IDs", group_label_size = 4, cell_size = 0.5,show_trajectory_graph = T)
```

##Step 3: Cluster the cells
```{r}
# cds = cluster_cells(cds, resolution=0.02)
cds = cluster_cells(cds, resolution=0.004)
plot_cells(cds, cell_size = 0.5)

```

##Step 4: Learn a graph and order cells
```{r}
cds <- learn_graph(cds)
plot_cells(cds, cell_size = 0.5)
```
```{r}
# ## With Shiny
# cds <- order_cells(cds)
# #
# save(file = "cdsRoots.rda", cds) #Save object state after selection of roots to avoid changes in furure runs
# cds_sub0.1 <- choose_graph_segments(cds)
# save(file = "cds_sub0.1.rda", cds_sub0.1)
# 
# cds_sub2.0 <- choose_graph_segments(cds)
# save(file = "cds_sub2.0.rda", cds_sub2.0)

## Without Shiny
load("cdsRoots.rda") #Load object
load("cds_sub0.1.rda")
load("cds_sub2.0.rda")

print(cds@principal_graph_aux$UMAP$root_pr_nodes)


```
```{r}
plot_cells(cds)
plot_cells(cds_sub0.1)
plot_cells(cds_sub2.0)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "Cluster.IDs",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5)

plot_cells(cds_sub0.1,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5)

plot_cells(cds_sub2.0,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5)
```

##Plot the pseudotime graph without trajectory
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5,
           show_trajectory_graph = FALSE)
```

## Correlation Analysis
```{r,results='hide'}
cds_pg <- graph_test(cds, neighbor_graph="principal_graph", cores=4, verbose = F)

cds_genes <- cds_pg %>% 
  filter(q_value < 0.05) %>% 
  arrange(desc(morans_I)) %>% 
  select(gene_short_name)
```

```{r}
cds_genes$gene_short_name
```

```{r}
# Plot genes
list_genes <- unique(c("unc-129" , "hum-2", "unc-53", "del-1", "bnc-1", "inx-12"	, "lgc-36", "unc-77", "dbl-1", "slo-2", "acr-2", "ser-2","egl-5","egl-5", "glr-5", "srb-16", "twk-46", "mig-13", "glr-4", "unc-3" , "lin-39", "mab-5", "unc-4", "vab-7", "unc-55", "cfi-1","mab-9", "exp-1", "lad-2", "lim-4", "lin-11", "avr-15", "ace-2", "ace-3","ace-4", "unc-86", "cho-1","cha-1", "unc-17", "unc-25", "glr-4", "glr-5","unc-129", "unc-3", "unc-77", "unc-17", "slo-2", "acr-2", "acr-16", "acr-5", "rig-4", "del-1","bnc-1","lgc-36", "eat-4","unc-30", "glr-4", "glr-5","acr-16", "acr-5", "rig-4", "unc-129", "del-1", "bnc-1", "lgc-36", "unc-8", "unc-53", "acr-2", "unc-17", "exp-1"))

nplots <- 9
x <- seq_along(list_genes)
toplot <- split(list_genes, ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
if (fillplot != 0) {
  toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])
}


for (x in toplot) {
  print(plot_cells(cds, genes= x,
             show_trajectory_graph=FALSE,
             label_cell_groups=FALSE,
             label_leaves=FALSE,
             cell_size = .4
            ))
}

```

##Pseudotemporal Expression Pattern
###Diferentially expressed cds_sub0.1
```{r}
nplots <- 4
x <- seq_along(as.character(cds_genes$gene_short_name[1:40]))
toplot <- split(as.character(cds_genes$gene_short_name[1:40]), ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
if (fillplot != 0) {
  toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])
}


for (x in toplot) {
  cds_lineage_cds <- cds_sub0.1[rowData(cds_sub0.1)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```

###Diferentially expressed cds_sub2.0
```{r}
nplots <- 4
x <- seq_along(as.character(cds_genes$gene_short_name[1:40]))
toplot <- split(as.character(cds_genes$gene_short_name[1:40]), ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
if (fillplot != 0) {
  toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])
}


for (x in toplot) {
  cds_lineage_cds <- cds_sub2.0[rowData(cds_sub2.0)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```

### Selected genes cds_sub0.1
```{r}

nplots <- 4
x <- seq_along(list_genes)
toplot <- split(list_genes, ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])



for (x in toplot) {
  cds_lineage_cds <- cds_sub0.1[rowData(cds_sub0.1)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```

### Selected genes cds_sub2.0
```{r}

nplots <- 4
x <- seq_along(list_genes)
toplot <- split(list_genes, ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])



for (x in toplot) {
  cds_lineage_cds <- cds_sub2.0[rowData(cds_sub2.0)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```
