---
title: "Pseudotime for clusters 42.1 and 56"
author: "Ramiro Lorenzo"
date: "April 8, 2020"
output:
  pdf_document: default
  html_document: default
---


## Load library
```{r library,results='hide',message=FALSE}
library(monocle3,verbose = FALSE)
library(dplyr,verbose = FALSE)
```

## Read the data
```{r Data}
## Loads a sparse matrix RawCountsPseudotime
load("../RawCountsPseudotime.rda")
dim(RawCountsPseudotime)
```

##Load cluster data
```{r}
## Loads a factor variabele ClusterPseudotime containing cluster identities
load("../ClusterPseudotime.rda")
## Change the name of variable to remind
DataClusters <- ClusterPseudotime

table(DataClusters)
length(DataClusters)
```

##Convert DataClusters to a matrix format for input to Monocle
```{r}
DataCluster.ID <- matrix(as.character(levels(DataClusters))[DataClusters], ncol = 1)
rownames(DataCluster.ID) <- names(DataClusters)
colnames(DataCluster.ID) <- "Cluster.IDs"
DataCluster.ID[1:10,]
```

##Generate matrix of gene short names for Monocle
```{r}
geneNames <- matrix(rownames(RawCountsPseudotime), ncol = 1)
rownames(geneNames) <- rownames(RawCountsPseudotime)
colnames(geneNames) <- "gene_short_name"
head(geneNames)
```

##Initiate Monocle object
```{r}
cds <- new_cell_data_set(expression_data = RawCountsPseudotime,
                         cell_metadata = DataCluster.ID,
                         gene_metadata = geneNames)

cds <- cds[,names(ClusterPseudotime[ClusterPseudotime %in% c(
                                                                # "0.0",
                                                                # "0.1",
                                                                # "1.0",
                                                                # "1.1",
                                                                # "2.0",
                                                                # "3.0",
                                                                # "3.1",
                                                                # "4.0",
                                                                # "4.1",
                                                                # "16.0",
                                                                # "16.1",
                                                                # "17.0",
                                                                # "17.1",
                                                                # "8.0.0",
                                                                # "8.0.1",
                                                                # "8.1",
                                                                # "8.2",
                                                                # "8.3",
                                                                # "10.0.0",
                                                                # "10.0.1_4",
                                                                # "10.0.2_3",
                                                                # "10.0.5",
                                                                # "10.1.0",
                                                                # "10.1.1",
                                                                # "10.2",
                                                                # "33.0",
                                                                # "42.0",
                                                                "42.1",
                                                                "56.0"
                                                                )])] #all data for trajectories



colData(cds)

## Column Cluster.IDs contains the original DataCluster IDs
colData(cds)$Cluster.IDs <- factor(colData(cds)$Cluster.IDs)
colData(cds)
```

## Step 1: Normalize and pre-process the data
```{r}
cds <- preprocess_cds(cds, num_dim = 40)
plot_pc_variance_explained(cds)
cds <- preprocess_cds(cds, num_dim = 10)
```

##Step 2: Reduce the dimensions using UMAP
```{r}
cds <- reduce_dimension(cds,umap.min_dist = 0.1,cores = 8)
plot_cells(cds, color_cells_by = "Cluster.IDs", group_label_size = 4, cell_size = 0.5,show_trajectory_graph = T)
```

##Step 3: Cluster the cells
```{r}
# cds = cluster_cells(cds, resolution=0.02)
cds = cluster_cells(cds, resolution=0.2)
plot_cells(cds, cell_size = 0.5)

```

##Step 4: Learn a graph and order cells
```{r}
cds <- learn_graph(cds)
plot_cells(cds, cell_size = 0.5)
```
```{r}
# ## With Shiny
# cds <- order_cells(cds)
# #
# save(file = "cdsRoots.rda", cds) #Save object state after selection of roots to avoid changes in furure runs

## Without Shiny
load("cdsRoots.rda") #Load object

print(cds@principal_graph_aux$UMAP$root_pr_nodes)
```
```{r}
plot_cells(cds)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "Cluster.IDs",
           label_groups_by_cluster=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE)
```

##Plot the pseudotime graph
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5)
```

##Plot the pseudotime graph without trajectory
```{r}
plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE, 
           graph_label_size=1.5,
           show_trajectory_graph = FALSE)
```

## Correlation Analysis
```{r,results='hide'}
cds_pg <- graph_test(cds, neighbor_graph="principal_graph", cores=4, verbose = F)

cds_genes <- cds_pg %>% 
  filter(q_value < 0.05) %>% 
  arrange(desc(morans_I)) %>% 
  select(gene_short_name)
```

```{r}
cds_genes$gene_short_name
```
```{r}
# Plot a few genes
list_genes <- unique(c("mpz-1","nrx-1", "unc-5", "unc-33","ina-1", "sel-10","mps-3","alr-1", "nhr-25", "tlp-1", "psa-3","lin-44","pdf-1", "vab-15","egl-20", "egl-27", "ceh-6", "egl-5", "jmjd-3.1", "che-14", "egl-26", "ajm-1","lin-26", "egl-26", "sem-4", "sox-2", "pdf-1", "ceh-9" , "mab-9","set-2", "cog-1", "exp-1", "tag-80", "madd-4","ina-1", "sel-10", "alr-1", "nhr-25", "tlp-1", "mps-3","cki-1","unc-3", "cho-1", "cha-1", "unc-17","sad-1", "syd-2", "unc-104", "unc-10", "egl-3", "egl-21"))

nplots <- 9
x <- seq_along(list_genes)
toplot <- split(list_genes, ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])


for (x in toplot) {
  print(plot_cells(cds, genes= x,
             show_trajectory_graph=FALSE,
             label_cell_groups=FALSE,
             label_leaves=FALSE,
             cell_size = .4
            ))
}

```

##Pseudotemporal Expression Pattern
###Diferentially expressed
```{r}
nplots <- 4
x <- seq_along(as.character(cds_genes$gene_short_name))
toplot <- split(as.character(cds_genes$gene_short_name), ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])



for (x in toplot) {
  cds_lineage_cds <- cds[rowData(cds)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```

### Selected genes
```{r}

nplots <- 4
x <- seq_along(list_genes)
toplot <- split(list_genes, ceiling(x/nplots))
fillplot <- nplots-length(toplot[[length(toplot)]])
toplot[[length(toplot)]] <- c(toplot[[length(toplot)]],toplot[[1]][1:fillplot])



for (x in toplot) {
  cds_lineage_cds <- cds[rowData(cds)$gene_short_name %in% x,]
  print(plot_genes_in_pseudotime(cds_lineage_cds,
                         min_expr=0.05))
}


```
