---
title: "ID Assignment and Cell Number"
author: "Jaun Ramiro Lorenzo"
date: "April 26, 2020"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


```{r libraries}
library(ggplot2)
library(sjmisc)
library(readr)
library(plyr)
library(readxl)

#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```


```{r load Dataset}
AverageExpression92.4 <- read_delim("../7-FinalPlotAndClusteringMarkers/AverageExpression92.4.csv", ",", escape_double = FALSE, trim_ws = TRUE)
ClusteringfinalAssignment <- read_excel("../Datasets/ClusteringfinalAssignment.xlsx")
cured_brain_atlas_PL_and_OH_CircularOtherMarkers <- read_delim("../Datasets/cured brain atlas PL and OH-CircularOtherMarkers.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
# load("../2-PlotBestFirstClustering/CeNeurons_Seurat_object_1st.rda")#seurat object
load("IterationData.rda")
IterationData <- unique(IterationData)

Clusters <- list("13" = c("CAN"), "14" = c("AVK"), "16" = c("FLP"), "30" = c("RIA"), "31" = c("ADL"), "34" = c("ASG"), "38" = c("ASJ"),"39" = c("RMG"), "40" = c("ASK"), "41" = c("DVA"), "44"= c("BAG"), "45"= c("AIN"), "46" = c("IL2"), "49" = c("AWA"), "52" = c("AVG"), "53" = c("RIC"), "55" = c("PVQ"), "59" = c("RIS"), "60" = c("M1"), "62" = c("ASER"), "63" = c("I5"))



IterationData <- IterationData[IterationData$parent_cluster %in% names(Clusters),]
IterationData <- cbind(IterationData, Good_assignment = NA)
IterationData <- cbind(IterationData, Perfect_assignment = NA)

sum_stats <- data.frame()
sum_stats_perfect <- data.frame()
#Check assignment
for (Cluster in names(Clusters)) {
  
  IterationData[IterationData$parent_cluster == Cluster,]$Good_assignment <- unlist(
    lapply(IterationData[IterationData$parent_cluster == Cluster,]$cluster_id,
           FUN = str_contains,pattern = Clusters[[Cluster]], 
           logic = "AND"))
  
  IterationData[IterationData$parent_cluster == Cluster,]$Perfect_assignment <- IterationData[IterationData$parent_cluster == Cluster,]$n_ident == length(Clusters[[Cluster]]) & IterationData[IterationData$parent_cluster == Cluster,]$Good_assignment
  
  length(Clusters[[Cluster]])
  
  ClusterData <- IterationData[IterationData$parent_cluster == Cluster,]
  
  #save statistics
  sum_stats <- rbind(sum_stats , 
                     data.frame(
                       c(Cluster = Cluster,
                         Good_assignment = T,
                         as.list(summary(ClusterData[ClusterData$Good_assignment == T,
                                                     c("n_cells")]))
                         )))
  
  sum_stats <- rbind(sum_stats , 
                     data.frame(
                       c(Cluster = Cluster,
                         Good_assignment = F,
                         as.list(summary(ClusterData[ClusterData$Good_assignment == F,
                                                     c("n_cells")]))
                         )))
  
  sum_stats_perfect <- rbind(sum_stats_perfect , 
                     data.frame(
                       c(Cluster = Cluster,
                         Perfect_assignment = T,
                         as.list(summary(ClusterData[ClusterData$Perfect_assignment == T,
                                                     c("n_cells")]))
                         )))
  
    sum_stats_perfect <- rbind(sum_stats_perfect , 
                     data.frame(
                       c(Cluster = Cluster,
                         Perfect_assignment = F,
                         as.list(summary(ClusterData[ClusterData$Perfect_assignment == F,
                                                     c("n_cells")]))
                         )))
  
}


#Assign Ranges
max(IterationData$n_cells)
min(IterationData$n_cells)

IterationData <- cbind(IterationData, Cell_range = NA)
irange <- 5
i <- 0
ranges <- c()
for (cell_range in 0:((50-irange) / irange) * irange) {
  IterationData$Cell_range[IterationData$n_cells > cell_range & IterationData$n_cells <= cell_range+irange] <- paste(cell_range+1,cell_range+irange, sep = "-")
  ranges <- c(ranges,paste(cell_range+1,cell_range+irange, sep = "-"))
} 
  IterationData$Cell_range[IterationData$n_cells > 50] <- "51-..."
  ranges <- c(ranges,"51-...")


Data_ranges <- data.frame()
Data_ranges_clusters <- data.frame()
for (r in ranges) {
  Good_assignement = sum(IterationData$Cell_range == r & IterationData$Good_assignment == T)
  Bad_assignement = sum(IterationData$Cell_range == r & IterationData$Good_assignment == F)
  P_Good_assignement = sum(IterationData$Cell_range == r & IterationData$Perfect_assignment == T)
  P_Bad_assignement = sum(IterationData$Cell_range == r & IterationData$Perfect_assignment == F)
  
  Data_ranges <- rbind(Data_ranges,
                       data.frame(cell_range = r,
                                  Good_assignement,
                                  Bad_assignement,
                                  Percentage = Good_assignement/(Good_assignement+Bad_assignement),
                                  P_Good_assignement,
                                  P_Bad_assignement,
                                  P_Percentage = P_Good_assignement/(P_Good_assignement+P_Bad_assignement)
                         
                       )
                       )
  
  
  for (cluster in names(Clusters)) {
      Good_assignement = sum(IterationData$Cell_range == r & IterationData$Good_assignment == T & IterationData$parent_cluster == cluster)
      Bad_assignement = sum(IterationData$Cell_range == r & IterationData$Good_assignment == F &  IterationData$parent_cluster == cluster)
      P_Good_assignement = sum(IterationData$Cell_range == r & IterationData$Perfect_assignment == T & IterationData$parent_cluster == cluster)
      P_Bad_assignement = sum(IterationData$Cell_range == r & IterationData$Perfect_assignment == F & IterationData$parent_cluster == cluster)
      
      Data_ranges_clusters <- rbind(Data_ranges_clusters,
                           data.frame(cell_range = r,
                                      assign = Good_assignement,
                                      Percentage = Good_assignement / (Good_assignement+Bad_assignement),
                                      P_assign = P_Good_assignement,
                                      P_Percentage = P_Good_assignement / (P_Good_assignement+P_Bad_assignement),
                                      Cluster = paste(cluster,Clusters[cluster], sep = "_"),
                                      Good_assign = T
      
                         
                       )
                       )

      Data_ranges_clusters <- rbind(Data_ranges_clusters,
                           data.frame(cell_range = r,
                                      assign = Bad_assignement,
                                      Percentage = Bad_assignement / (Good_assignement+Bad_assignement),
                                      P_assign = P_Bad_assignement,
                                      P_Percentage = P_Bad_assignement / (P_Good_assignement+P_Bad_assignement),
                                      Cluster = paste(cluster,Clusters[cluster], sep = "_"),
                                      Good_assign = F))
    
    
    
  }
}
Data_ranges


```


# Summary statistics (Mean values for all clusters) - Iteration data from second screening where small clusters are obtained by increasing resolution
## Number of cells for sub-clusters when the automatically assigned identities include the expected ones
```{r}
#-Properly assigned:
apply(sum_stats[sum_stats$Good_assignment==T,c("Min.", "X1st.Qu.", "Median", "Mean", "X3rd.Qu.", "Max.")], 2, mean)
#-Badly assigned:
apply(sum_stats[sum_stats$Good_assignment==F,c("Min.", "X1st.Qu.", "Median", "Mean", "X3rd.Qu.", "Max.")], 2, mean)

```

## Number of cells for sub-clusters when the automatically assigned identities match exactly the expected
```{r}
#Two clusters do not contain perfect matches
sum_stats_perfect <- sum_stats_perfect[sum_stats_perfect$Cluster != "43",]
sum_stats_perfect <- sum_stats_perfect[sum_stats_perfect$Cluster != "48",]
#-Properly assigned:
apply(sum_stats_perfect[sum_stats_perfect$Perfect_assignment==T,c("Min.", "X1st.Qu.", "Median", "Mean", "X3rd.Qu.", "Max.")], 2, mean)
#-Badly assigned:
apply(sum_stats_perfect[sum_stats_perfect$Perfect_assignment==F,c("Min.", "X1st.Qu.", "Median", "Mean", "X3rd.Qu.", "Max.")], 2, mean)
```


# Boxplots
## Number of cells for sub-clusters when the automatically assigned identities include the expected ones
```{r fig.width=20}

# Box plot
p <- ggplot(IterationData, aes(x=parent_cluster, y=n_cells, fill = Good_assignment)) + ylim(0,150) + 
    theme(text = element_text(size=20)) +
  geom_boxplot()
p

# Box plot

p <- ggplot(Data_ranges_clusters, aes(x=cell_range, y=Percentage, fill = Good_assign)) +
    theme(text = element_text(size=20)) +
  geom_boxplot()
p

```

# Percenage of good assign per cell range
```{r}

dfsumamry <- data_summary(Data_ranges_clusters, varname="Percentage", 
                    groupnames=c("cell_range","Good_assign"))

# Line plot with good assignement
ggplot(dfsumamry, aes(x=cell_range, y=Percentage, group=Good_assign, color=Good_assign)) + 
    geom_errorbar(aes(ymin=Percentage-sd, ymax=Percentage+sd), width=.1) +
    geom_line() + geom_point()

dfsumamry <- data_summary(Data_ranges_clusters, varname="P_Percentage", 
                    groupnames=c("cell_range","Good_assign"))

# Line plot with perfect assignement
ggplot(dfsumamry, aes(x=cell_range, y=P_Percentage, group=Good_assign, color=Good_assign)) + 
    geom_errorbar(aes(ymin=P_Percentage-sd, ymax=P_Percentage+sd), width=.1) +
    geom_line() + geom_point()

```


## Number of cells for sub-clusters when automatically assigned identities match exactly the expected 
```{r fig.width=20}

# Box plot
p <- ggplot(IterationData, aes(x=parent_cluster, y=n_cells, fill = Perfect_assignment)) + ylim(0,150) + 
    theme(text = element_text(size=20)) +
  geom_boxplot()
p

```

#Statistics for Clusters automaticaly assigned in the first round (Final data from clustering parameters PCs = 92 and Resol = 4)

## Gene atlas summary
```{r}
cured_brain_atlas_PL_and_OH_CircularOtherMarkers <- cured_brain_atlas_PL_and_OH_CircularOtherMarkers[,2:length(cured_brain_atlas_PL_and_OH_CircularOtherMarkers)]
nGene <- apply(cured_brain_atlas_PL_and_OH_CircularOtherMarkers[,unlist(Clusters)], 2, sum)
names(nGene) <- paste(unlist(names(Clusters)),names(nGene),sep = "_")
print("Total atlas genes per neuron class")
nGene

specificGenes <- apply(cured_brain_atlas_PL_and_OH_CircularOtherMarkers, 1, sum) == 1
cured_brain_atlas_specific <- cured_brain_atlas_PL_and_OH_CircularOtherMarkers[specificGenes,]
nGene <- apply(cured_brain_atlas_specific[,unlist(Clusters)], 2, sum)
names(nGene) <- paste(unlist(names(Clusters)),names(nGene),sep = "_")
print("Specific atlas genes (only expresed in one neuron)")
nGene

```

# Total genes detected vs cluster size (N_cells)
```{r}
ClusteringfinalAssignment <- as.data.frame(ClusteringfinalAssignment)
rownames(ClusteringfinalAssignment) <- ClusteringfinalAssignment$Cluster
totalGenes <- apply(AverageExpression92.4>0, 2, sum)
totalGenes <- totalGenes[3:length(totalGenes)-1]

dataGeneCount <- data.frame(N_cells = ClusteringfinalAssignment[names(totalGenes),"N Cells"],Total_gene_count = totalGenes)

ClusteringfinalAssignment[names(totalGenes),"N Cells"]
ggplot(dataGeneCount, aes(N_cells,Total_gene_count))+
  geom_point() +
  geom_smooth(method = "lm",formula=y~log(x))

#log(N_cells)
ggplot(dataGeneCount, aes(log(N_cells),Total_gene_count))+
  geom_point() +
  geom_smooth(method = "lm")



cor(log(dataGeneCount$N_cells),dataGeneCount$Total_gene_count)

```

